---
title: <center>AMARETTO Regulatory Module `r toString(ModuleNr)` Report</center>
pagetitle: module `r toString(ModuleNr)` report
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cerulean
    highlight: NULL
---

<style>
  .main-container {
    max-width:  100% !important;
    padding-right: 1% !important;
    padding-left: 1% !important;
    margin-right: 1% !important;
    margin-left: 1% !important;
  }
</style>

<center>
<br><br>
<a href="../index.html"><button type="button"  class="btn btn-primary">Index Page</button></a>
<br><br>

## Regulatory Module Heatmap Visualization
<br><br><br>
```{r moduleheatmap, echo=FALSE, fig.height = 12, fig.width = 16, fig.align = "center"}
suppressMessages(library(ComplexHeatmap))
ComplexHeatmap::draw(heatmap_module,heatmap_legend_side = "bottom",annotation_legend_side="bottom")
```
<br><br><br>

## Activator and Repressor Driver Genes in Regulatory Module
<br><br><br>
```{r regulators, echo=FALSE}
dt_regulators
```
<br><br><br>

## Target Genes in Regulatory Module
<br><br><br>
```{r targets, echo=FALSE}
dt_targets
```
<br><br><br>

## Enrichments of Functional Categories in Regulatory Module
<br><br><br>
```{r geneset, echo=FALSE, out.width='25%'}
dt_genesets
```

<br><br><br>

## Associations of Phenotypes to Regulatory Module

<br><br><br>

```{r module_Phenotypes, echo=FALSE}
dt_phenotype_association
```

<br><br><br>

</center>
