---
title: <center>AMARETTO Report</center>
pagetitle: AMARETTO Report
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cerulean
    highlight: NULL
---

<style>
  .main-container {
    max-width:  100% !important;
    padding-right: 1% !important;
    padding-left: 1% !important;
    margin-right: 1% !important;
    margin-left: 1% !important;
    table-layout: fixed !important;
  }
</style>

<center>
## Run Information
<br>
Number of Samples in Gene Expression Data = `r nExp`  
Number of Samples in DNA Copy Number Data = `r nCNV`  
Number of Samples in DNA Methylation Data = `r nMET`  
Number of `r VarPercentage`% most variable Genes = `r nGenes`  
Number of Regulatory Modules = `r nMod`  
<br><br>

## Overview of Regulatory Modules
<br><br><br>
```{r Modules,echo=FALSE}
dt_overview
```
<br><br><br>

## Assignments of Genes to Regulatory Modules
<br><br><br>
```{r Genes,echo=FALSE}
dt_genes
```
<br><br><br>

## Enrichments of Functional Categories in Regulatory Modules
<br><br><br>
```{r Genesets,echo=FALSE}
dt_genesetsall
```

<br><br><br>

## Associations of Phenotypes to Regulatory Modules

<br><br><br>

```{r Phenotypes,echo=FALSE}
dt_phenotype_association_all
```
<br><br><br>

</center>
