---
title: <center>AMARETTO Report</center>
output: 
  html_document:
    theme: cerulean
    highlight: NULL
---
<meta name="viewport" content="width=device-width, initial-scale=1.00, user-scalable=yes">
<center>
## Run Information
<br>
Number of Samples in Gene Expression Data = `r nExp`  
Number of Samples in DNA Copy Number Data = `r nCNV`  
Number of Samples in DNA Methylation Data = `r nMET`  
Number of `r VarPercentage`% most variable Genes = `r nGenes`  
Number of Regulatory Modules = `r nMod`  
<br><br>

## Overview of Regulatory Modules
<br><br><br>
```{r Modules,echo=FALSE}
dt_overview
```
<br><br><br>

## Assignments of Genes to Regulatory Modules
<br><br><br>
```{r Genes,echo=FALSE}
dt_genes
```
<br><br><br>

## Enrichments of Functional Categories in Regulatory Modules
<br><br><br>
```{r Genesets,echo=FALSE}
dt_genesetsall
```

<br><br><br>

## Associations of Phenotypes to Regulatory Modules

<br><br><br>

```{r Phenotypes,echo=FALSE}
phenotype_association_datatable
```
<br><br><br>

</center>
